---
title: "Class 11: Genome Informatics"
author: "Winnie Zhou (A16673200)"
format: pdf
---


# Section 1. Proportion og G/G in a population

Here we read the CSV file that we downloaded from Ensemble

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl) * 100
```

Now let's look at a different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.)/nrow(gbr)*100)
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MXL population.

Lets now dig into this further.


#Section 4: Population Scale Analysis
One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 


How many samples do we have

```{r}
expr <- read.table(url("https://bioboot.github.io/bimm143_S24/class-material/rs8067378_ENSG00000172057.6.txt"))
head(expr)
```

```{r}
nrow(expr)
```

How many of each type?

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```

Let's make a boxplot.

```{r}
ggplot(expr) + aes(x=geno, exp, fill=geno) +
  geom_boxplot(notch=TRUE)
```

From this plot, we can see the relative expression level of A/A is hgigher than the relative expression level of G/G. The SNP does seem to effect the expression of ORMDL3.








